---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ragnar

🚧 Under active development 🚧

<!-- badges: start -->
<!-- badges: end -->

`ragnar` is an R package that helps implement Retrieval-Augmented Generation (RAG) workflows. It focuses on providing a complete solution with sensible defaults, while still giving the knowledgeable user precise control over each steps. We don't believe that you can fully automate the creation of a good RAG system, so it's important that `ragnar` is not a black box. `ragnar` is designed to be transparent—you can inspect easily outputs at intermediate steps to understand what's happening.

## Installation

```r
pak::pak("t-kalinowski/ragnar")
```

## Key Steps

### 1. Document Processing

`ragnar` works with a wide variety of document types, using [MarkItDown](https://github.com/microsoft/markitdown) to convert content to Markdown.

Key functions:

- `ragnar_find_links()`: Find all links in a webpage
- `ragnar_read()`: Convert a file or URL to markdown

### 2. Text Chunking

Next we divide each document into multiple chunks. Ragnar defaults to a strategy that preserves some of the semantics of the document, but provide plenty of options to tweak the approach.

Key functions:

- `ragnar_chunk()`: Higher-level function that both identifies semantic boundaries and chunks text.
- `ragnar_segment()`: Lower-level function that identifies semantic boundaries.
- `ragnar_chunk_segments()`: Lower-level function that chunks pre-segmented text.

### 3. Context Augmentation (Optional)

RAG applications benefit from augmenting text chunks with additional context, such as document headings and subheadings. While `ragnar` doesn't directly export functions for this, it supports template-based augmentation through `ragnar_read(frame_by_tags, split_by_tags)`. Future versions will support generating context summaries via LLM calls.

Key functions:

- `ragnar_read()`: Use `frame_by_tags` and/or `split_by_tags` arguments to associate text chunks with their document position.
- `markdown_segment()`: Segment markdown text into a character vector using semantic tags (e.g., headings, paragraphs, or code chunks).
- `markdown_frame()`: Convert markdown text into a dataframe.

### 4. Embedding

`ragnar` can help compute embeddings for each chunk. The goal is for `ragnar` to provide access to embeddings from popular LLM providers. Currently only `ollama` and `openai` providers.

Key functions:

- `embed_ollama()`
- `embed_openai()`

### 5. Storage

Processed data is stored in a format optimized for efficient searching, using `duckdb` by default. The API is designed to be extensible, allowing additional packages to implement support for different storage providers.

Key functions:

- `ragnar_store_create()`
- `ragnar_store_connect()`
- `ragnar_store_insert()`

### 6. Retrieval

Given a prompt, retrieve related chunks based on embedding distance or bm25 text search.

Key functions:

- `ragnar_retrieve()`
- `ragnar_retrieve_vss()`: Retrieve using [`vss` DuckDB extension](https://duckdb.org/docs/extensions/vss.html)
- `ragnar_retrieve_bm25()`: Retrieve using [`full-text search DuckDB extension`](https://duckdb.org/docs/extensions/full_text_search.html)

### 7. Re-ranking (Optional)

Re-ranking of retrieved chunks is planned for future releases.

### 8. Prompt Generation

`ragnar` will provide tools for incorporating retrieved chunks into LLM prompts.

## Usage

Here's an example of using `ragnar` to store and retrieve chunks from the *R for Data Science (2e)* book:

```{r, code = readLines("examples/example.R")}
```
