---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ragnar

🚧 Under active development 🚧

<!-- badges: start -->
<!-- badges: end -->

`ragnar` is an R package that helps implement Retrieval-Augmented Generation (RAG) workflows. It focuses on providing a complete solution with sensible defaults, while still giving the knowledgeable user precise control over all the steps. We don't believe that you can fully automate the creation of a good RAG, so it's important that ragnar is not a black box; `ragnar` is designed to be transparent—you can inspect outputs at intermediate steps to understand what’s happening.


## Installation
```r
pak::pak("t-kalinowski/ragnar")
```

## Key Steps

### 1. Document Processing

Ragnar starts with a directory of markdown or HTML files. In the long term, we plan to offer tools for building this directory, such as [web scraping](https://github.com/r-lib/httr2/pull/584) or markdown processing. We also intend to add tools for simplifying HTML using readability techniques or a pandoc AST walker to remove extraneous attributes.

### 2. Text Chunking

Next we divide each document into multiple chunks. We default to a strategy that preserves some of the semantics of the document, but provide plenty of options to tweak the approach.

### 3. (Optionally) Context augmentation

Add needed context to the chunk. The goal it to support a variety of workflows for attaching context to chunked text, ranging from
templating strings `glue()` to sophisticated workflows that involve calling out to LLMs to generate relevant contextualized summaries.

### 4. Embedding

Support computing an embedding for each chunk. The goal is for `ragnar` to provide access to embeddings from popular LLM providers. Currently, only support for `ollama` is implemented.

### 5. Storage

Store all processed data in a format optimized for efficient searching, using `duckdb` as a default.
The goal is for the API to be extensible, so additional packages to implement support for different store providers.

### 6. Retrieval

Retrieve related chunks based on cosine similarity of embeddings. In the near future we intend also support retrieval using BM25 ranking and regular text search.

- **Vector similarity search:** [`vss` DuckDB extension](https://duckdb.org/docs/extensions/vss.html)
- **Text search:** [`full-text search extension`](https://duckdb.org/docs/extensions/full_text_search.html)

### 7. (Optional) Re-ranking

Rerank retreived chunks. (Not yet implemented)

### 8. Prompt generation

Ragnar will provide a set of tools for helping to incorporate the retrieved chunks into an LLM prompt.


## Usage

Below is an example of using `ragnar` to store and retrieve chunks from the *R for Data Science (2e)* book.


```{r, code = readLines("examples/example.R")}
```

<!-- ```{r, eval = FALSE} -->
<!-- # Create a local copy of the rendered book -->
<!-- if (!dir.exists("~/github/hadley/r4ds")) { -->
<!--   # Clone https://r4ds.hadley.nz/base-r locally -->
<!--   system2("git", c("clone", -->
<!--     shQuote("https://github.com/hadley/r4ds/"), -->
<!--     shQuote(normalizePath("~/github/hadley/r4ds")) -->
<!--   )) -->
<!--   pak::local_install_dev_deps("~/github/hadley/r4ds") -->
<!--   system("quarto render ~/github/hadley/r4ds") -->
<!-- } -->
<!-- ``` -->
